An exposition on the propriety of restricted Boltzmann machines

### 2016-09-17 writing/appendix.Rmd
formatting

For integers $m,n \geq 1$, consider random vectors $\bm{V}_m \equiv(V_1,\ldots,V_m)$ and $\bm{H}_n\equiv (H_1,\ldots,H_n)$, where the random variables $V_i$ and $H_j$ assume values in $\{1,-1\}$, $i=1,\ldots,m$, $j=1,\ldots,n$. We suppose the vector $(\bm{V}_m,\bm{H}_n)$ follows a restricted Boltzmann machine (RBM) distribution

\begin{eqnarray*}
&&P(V_1= v_1,\ldots,V_m= v_m, H_1= h_1,\ldots,H_n=h_n|\bm{\theta})   \\&\propto &\exp\left[ \sum_{i=1}^m v_i \theta_{v_j} + \sum_{j=1}^n h_j \theta_{h_j} + \sum_{i=1}^m  \sum_{j=1}^n v_i  h_j \theta_{ij} \right], \quad \{v_i\}_{i=1}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\},
\end{eqnarray*}
with $m+n+mn$ real parameters $\theta_{v_1},\ldots,\theta_{v_m},\theta_{h_1},\ldots,\theta_{h_m}$ and
$\theta_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$ collected in a parameter vector denoted as $\bm{\theta}\in \mathbb{R}^{m+n+mn}$.
Denote the distribution of $(\bm{V}_m,\bm{H}_n)$ as RBM$(m,n,\bm{\theta})$. In the following, for convenience, we say a symmetric random variable $X$ has a Bernoulli$(1/2)$ distribution if $P(X=1)=1/2=P(X=-1)$.

Denote the random vector $\bm{V}_m*\bm{H}_n = (V_1 H_1,\ldots,V_1 H_n, V_2 H_1,\ldots,V_2 H_n,\ldots, V_m H_1,\ldots,V_m H_n)$.
For the RBM$(m,n,\bm{\theta})$ model, we are interested in the function $\bm{g}: \mathbb{R}^{m+n+mn} \rightarrow   \mathcal{I} \subset \mathbb{R}^{m+n+mn}$ given by
\begin{eqnarray*}
\bm{g}(\bm{\theta}) &=& \E_{\bm{\theta}}(\bm{V}_m,\bm{H}_n,\bm{V}_m*\bm{H}_n)\\
& = & \sum_{\bm{v}_m \in \{\pm 1\}^m, \bm{h}_n \in \{\pm 1\}^n} (\bm{v}_m,\bm{h}_n,\bm{v}_m*\bm{h}_n)P(\bm{V}_m=\bm{v}_m,\bm{H}_n=\bm{h}_n|\bm{\theta})
\end{eqnarray*}
for $\bm{\theta}\in \mathbb{R}^{m+n+mn}$; above $\mathcal{I} \equiv \{\bm{g}(\bm{\theta}) : \bm{\theta}\in \mathbb{R}^{m+n+mn}\}$ denotes the image of the mapping.

@@

**Proof of Theorem 1.** We first establish the equivalence of claims 2 and 3. If claim 2 holds, then
 claim 3 holds trivially. Suppose claim 3 holds and pick $\ell \in \{1,\ldots,m\}$. Let  $\theta_{v_1},\ldots,\theta_{v_m},\theta_{h_1},\ldots,\theta_{h_n}$, $\theta_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the components of $\bm{\theta}_1$ and let $\tilde{\theta}_{v_1},\ldots,\tilde{\theta}_{v_m},\tilde{\theta}_{h_1},\ldots,\tilde{\theta}_{h_n}$, $\tilde{\theta}_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the analogous components of $\bm{\theta}_2$. Then, by assumption,
\begin{eqnarray*}
&& \exp\left[ 2\theta_{v_\ell} +2 \sum_{j=1}^n  h_j \theta_{\ell j} \right]\\
&=&  \frac{P_1\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)} {P_1\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\\
&=& \frac{P_2\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}
{P_2\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\\
&& \exp\left[ 2\tilde{\theta}_{v_\ell} +2 \sum_{j=1}^n  h_j \tilde{\theta}_{\ell j} \right]
\end{eqnarray*}
for any $\{v_i\}_{i=1,i\neq \ell}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\}$. Equating exponents, we have $\theta_{v_\ell} - \tilde{\theta}_{v_\ell} = \sum_{j=1}^n  h_j (  \tilde{\theta}_{\ell j} -\theta{\ell j})$ for any $\{h_j\}_{j=1}^n \subset \{\pm 1\}$. Taking all $h_j=1$ or all $h_j=-1$ for $1 \leq j \leq n$ yields $\theta_{v_\ell} = \tilde{\theta}_{v_\ell}$. This result further implies $\tilde{\theta}_{\ell 1} -\theta{\ell 1} = -  \sum_{j=2}^n  h_j (  \tilde{\theta}_{\ell j} -\theta_{\ell j})$ for
any given $\{h_j\}_{j=2}^n \subset \{\pm 1\}$, so that $\theta{\ell 1}=\tilde{\theta}_{\ell 1}$ must hold. Iterating this argument, sequentially for each $j=2,\ldots,m$, shows $\theta{\ell j}=\tilde{\theta}_{\ell j}$ for any $1 \leq j \leq n$ in addition to $\theta_{v_\ell}=\tilde{\theta}_{v_\ell}$.  As $\ell\in\{1,\ldots,m\}$ was arbitrary, we have $\theta_{v_i}=\tilde{\theta}_{v_i}$ and $\theta{i j}=\tilde{\theta}_{i j}$ for any $1 \leq i \leq m$, $1 \leq j \leq n$.

@@

Considering random variables $(\bm{V}_{m+1},\bm{H}_n) = (V_1,\ldots,V_m,V_{m+1}, H_1,\ldots,H_n)$, the induction hypothesis
applies to $(\bm{V}_{m},\bm{H}_n)$ where $\bm{V}_m =(V_1,\ldots,V_m)$. Hence, for any $\bm{v}_m \in \{a,b\}^{m}, \bm{h}_n \in \{a,b\}^{n}$, we have $P_1(\bm{V}_m=\bm{v}_m,\bm{H}_n=\bm{h}_n) = P_2(\bm{V}_m=\bm{v}_m,\bm{H}_n=\bm{h}_n)>0$, $P_1( \bm{H}_n=\bm{h}_n) = P_2(\bm{H}_n=\bm{h}_n)>0$ and consequently
\begin{equation}\label{eqn0}
P_1(\bm{V}_m=\bm{v}_m|\bm{H}_n=\bm{h}_n) =P_2(\bm{V}_m=\bm{v}_m |\bm{H}_n=\bm{h}_n).
\end{equation}
Likewise, by the induction hypothesis applied to $(\bm{V}_{m-},\bm{H}_n)$ for $\bm{V}_{m-} = (V_2,\ldots,V_{m+1})$, we have
\begin{eqnarray}\label{eqn1}
 1- P_1(V_{m+1} = b | \bm{H}_n=\bm{h}_n) &=& P_1(V_{m+1} = a |\bm{H}_n=\bm{h}_n)\\
\nonumber  &=&  P_2(V_{m+1} = a | \bm{H}_n=\bm{h}_n)\\
\nonumber  & = &1-P_2(V_{m+1} = b | \bm{H}_n=\bm{h}_n)
\end{eqnarray}
for any $\bm{h}_n \in \{a,b\}^{n}$. Now fix  $\bm{v}_m \in \{a,b\}^{m}, \bm{h}_n \in \{a,b\}^{n}$ and let $\bm{v}_{m+1}= (v_{m+1},\bm{v}_m)$ for $v_{m+1}\in\{a,b\}$.  Then, by the conditional independence assumption,
$$
P_i( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n) = P_i( V_{m+1}=v_{m+1}| \bm{H}_n=\bm{h}_n) P_i( \bm{V}_{m}=v_{m}| \bm{H}_n=\bm{h}_n)
$$
for $i=1,2,$,  so that we have   $P_1( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n) = P_2( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n)$ by (\ref{eqn0})-(\ref{eqn1}). Consequently, it follows that
\begin{eqnarray*}
P_1( \bm{V}_{m+1}=\bm{v}_{m+1}, \bm{H}_n=\bm{h}_n) &=& P_1( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n) P_1(\bm{H}_n=\bm{h}_n)\\
& =& P_2( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n)P_2(\bm{H}_n=\bm{h}_n) \\&=&P_2( \bm{V}_{m+1}=\bm{v}_{m+1}, \bm{H}_n=\bm{h}_n)
\end{eqnarray*}
by $P_1(\bm{H}_n=\bm{h}_n)=P_2(\bm{H}_n=\bm{h}_n)$. As $\bm{v}_{m+1}\in \{a,b\}^{m+1},\bm{h}_n \in \{a,b\}^{n}$ were arbitrary, this completes the proof of Lemma 1. $\Box$

\begin{lemma} Suppose the discrete random vector $(X,Y)$ has support $\{a,b\}\times \{a,b\}$, for some $a\neq b \in \mathbb{R}$,
under two probability distributions $P_1$ and $P_2$. Let $\E_i$ denote expectation under the distribution $P_i$, $i=1,2$. Then, the following are equivalent
\begin{enumerate}
\item $\E_1 X = \E_2 X$, $\E_1 Y = \E_2 Y$, $ \E_1 X Y = \E_2 X Y$, where $\E_i$ denotes expectation under $P_i$, $i=1,2$.
\item $P_1(X=x,Y=y)=P_2(X=x,Y=y)$ for $x,y\in\{a,b\}$, i.e., $(X,Y)$ has the same distribution under $P_1$ and $P_2$.
\end{enumerate}
\end{lemma}

*Proof.* Denote the four probabilities by which the random vector $(X,Y)$ assume pairs $(b,b),(b,a),(a,b),(a,a)$, respectively, as $c_1,c_2,c_3,1-c_1-c_2-c_3$ under $P_1$ and $d_1,d_2,d_3,1-d_1-d_2-d_3$ under $P_2$.  The conditions $\E_1 X = \E_2 X$, $\E_1 Y = \E_2 Y$, $ \E_1 X Y = \E_2 X Y$ with $b\neq a$ imply that
$$
(c_1+c_2) = (d_1+d_2), \quad  (c_1+c_3) = (d_1+d_3) \quad (b-a) c_1 + a (2c_1+c_2+c_3) =  (b-a) d_1 + a (2d_1+d_2+d_3).
$$
As $(2c_1+c_2+c_3) = (2d_1+d_2+d_3) $ and $b\neq a$, we conclude that $c_1=d_1$, from which it follows that $c_2=d_2$ and $c_3=d_3$. Hence, claim 1 implies claim 2 in Lemma 2.  Claim 2 also trivially implies claim 1. $\Box$

@@

Now suppose claim 3 holds. Pick $\ell \in \{1,\ldots,m\}$. Then, by the RBM$(m,n,\bm{\theta})$ probability structure and the iid Bernoulli assumption, we have the conditional probability
\begin{eqnarray*}
\frac{1}{2} &=& P\left(V_\ell = v_{\ell}\Big| \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right) \\
&\equiv &\frac{\exp\left( v_\ell\theta_\ell +  v_\ell\sum_{j=1}^n  h_j \lambda_{\ell,j}\right)}{\exp\left(
\theta_\ell + \sum_{j=1}^n   h_j\lambda_{\ell,j} \right) +\exp\left(-\theta_\ell  -\sum_{j=1}^n  h_j\lambda_{\ell,j}\right) }
\end{eqnarray*}
for any subsets $\{v_i\}_{i=1}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\}$. Equating conditional probabilities when $v_\ell=1$ or when $v_\ell=-1$, we have   $\theta_\ell = -  \sum_{j=1}^m  h_j \lambda_{\ell,j}$ for any given $\{h_j\}_{j=1}^n \subset \{\pm 1\}$. Taking all $h_j=1$ or all $h_j=-1$ for $1 \leq j \leq n$ yields $\theta_\ell=0$.  This result further implies $\lambda_{\ell,1} = - \sum_{j=2}^n  h_j \lambda_{\ell,j}$ for any given $\{h_j\}_{j=2}^n \subset \{\pm 1\}$, so that $\lambda_{\ell,1}=0$ must hold. Iterating this argument, sequentially for each $j=2,\ldots,m$, shows $\lambda_{\ell,j}=0$ for any $1 \leq j \leq n$ in addition to $\theta_\ell=0$. As $\ell\in\{1,\ldots,m\}$ was arbitrary, we have $\theta_i=0$ and $\lambda_{i,j}=0$ for any $1 \leq i \leq m$, $1 \leq j \leq n$.

>>>>

For integers $m,n \geq 1$, consider random vectors $\bm{V}_m \equiv(V_1,\ldots,V_m)$ and $\bm{H}_n\equiv (H_1,\ldots,H_n)$, where the random variables $V_i$ and $H_j$ assume values in $\{1,-1\}$, $i=1,\ldots,m$, $j=1,\ldots,n$. We suppose the vector $(\bm{V}_m,\bm{H}_n)$ follows a restricted Boltzmann machine (RBM) distribution
\begin{align*}
&P(V_1= v_1,\ldots,V_m= v_m, H_1= h_1,\ldots,H_n=h_n|\bm{\theta}) \propto \exp\left[ \sum_{i=1}^m v_i \theta_{v_j} + \sum_{j=1}^n h_j \theta_{h_j} + \sum_{i=1}^m  \sum_{j=1}^n v_i  h_j \theta_{ij} \right], \\
& \quad \{v_i\}_{i=1}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\},
\end{align*}
with $m+n+mn$ real parameters $\theta_{v_1},\ldots,\theta_{v_m},\theta_{h_1},\ldots,\theta_{h_m}$ and
$\theta_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$ collected in a parameter vector denoted as $\bm{\theta}\in \mathbb{R}^{m+n+mn}$.
Denote the distribution of $(\bm{V}_m,\bm{H}_n)$ as RBM$(m,n,\bm{\theta})$. In the following, for convenience, we say a symmetric random variable $X$ has a Bernoulli$(1/2)$ distribution if $P(X=1)=1/2=P(X=-1)$.

Denote the random vector $\bm{V}_m*\bm{H}_n = (V_1 H_1,\ldots,V_1 H_n, V_2 H_1,\ldots,V_2 H_n,\ldots, V_m H_1,\ldots,V_m H_n)$.
For the RBM$(m,n,\bm{\theta})$ model, we are interested in the function $\bm{g}: \mathbb{R}^{m+n+mn} \rightarrow   \mathcal{I} \subset \mathbb{R}^{m+n+mn}$ given by
\begin{align*}
\bm{g}(\bm{\theta}) &= \E_{\bm{\theta}}(\bm{V}_m,\bm{H}_n,\bm{V}_m*\bm{H}_n)\\
&= \sum_{\bm{v}_m \in \{\pm 1\}^m, \bm{h}_n \in \{\pm 1\}^n} (\bm{v}_m,\bm{h}_n,\bm{v}_m*\bm{h}_n)P(\bm{V}_m=\bm{v}_m,\bm{H}_n=\bm{h}_n|\bm{\theta})
\end{align*}
for $\bm{\theta}\in \mathbb{R}^{m+n+mn}$; above $\mathcal{I} \equiv \{\bm{g}(\bm{\theta}) : \bm{\theta}\in \mathbb{R}^{m+n+mn}\}$ denotes the image of the mapping.

@@

**Proof of Theorem 1.** We first establish the equivalence of claims 2 and 3. If claim 2 holds, then
 claim 3 holds trivially. Suppose claim 3 holds and pick $\ell \in \{1,\ldots,m\}$. Let  $\theta_{v_1},\ldots,\theta_{v_m},\theta_{h_1},\ldots,\theta_{h_n}$, $\theta_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the components of $\bm{\theta}_1$ and let $\tilde{\theta}_{v_1},\ldots,\tilde{\theta}_{v_m},\tilde{\theta}_{h_1},\ldots,\tilde{\theta}_{h_n}$, $\tilde{\theta}_{ij}$, $i=1,\ldots,m$, $j=1,\ldots,n$, denote the analogous components of $\bm{\theta}_2$. Then, by assumption,
\begin{align*}
\exp\left[ 2\theta_{v_\ell} +2 \sum_{j=1}^n  h_j \theta_{\ell j} \right] &= \frac{P_1\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)} {P_1\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\\
&= \frac{P_2\left(\{V_\ell = 1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}
{P_2\left(\{V_\ell = -1\} \cap \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right)}\exp\left[ 2\tilde{\theta}_{v_\ell} +2 \sum_{j=1}^n  h_j \tilde{\theta}_{\ell j} \right]
\end{align*}
for any $\{v_i\}_{i=1,i\neq \ell}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\}$. Equating exponents, we have $\theta_{v_\ell} - \tilde{\theta}_{v_\ell} = \sum_{j=1}^n  h_j (  \tilde{\theta}_{\ell j} -\theta{\ell j})$ for any $\{h_j\}_{j=1}^n \subset \{\pm 1\}$. Taking all $h_j=1$ or all $h_j=-1$ for $1 \leq j \leq n$ yields $\theta_{v_\ell} = \tilde{\theta}_{v_\ell}$. This result further implies $\tilde{\theta}_{\ell 1} -\theta{\ell 1} = -  \sum_{j=2}^n  h_j (  \tilde{\theta}_{\ell j} -\theta_{\ell j})$ for
any given $\{h_j\}_{j=2}^n \subset \{\pm 1\}$, so that $\theta{\ell 1}=\tilde{\theta}_{\ell 1}$ must hold. Iterating this argument, sequentially for each $j=2,\ldots,m$, shows $\theta{\ell j}=\tilde{\theta}_{\ell j}$ for any $1 \leq j \leq n$ in addition to $\theta_{v_\ell}=\tilde{\theta}_{v_\ell}$.  As $\ell\in\{1,\ldots,m\}$ was arbitrary, we have $\theta_{v_i}=\tilde{\theta}_{v_i}$ and $\theta{i j}=\tilde{\theta}_{i j}$ for any $1 \leq i \leq m$, $1 \leq j \leq n$.

@@

Considering random variables $(\bm{V}_{m+1},\bm{H}_n) = (V_1,\ldots,V_m,V_{m+1}, H_1,\ldots,H_n)$, the induction hypothesis
applies to $(\bm{V}_{m},\bm{H}_n)$ where $\bm{V}_m =(V_1,\ldots,V_m)$. Hence, for any $\bm{v}_m \in \{a,b\}^{m}, \bm{h}_n \in \{a,b\}^{n}$, we have $P_1(\bm{V}_m=\bm{v}_m,\bm{H}_n=\bm{h}_n) = P_2(\bm{V}_m=\bm{v}_m,\bm{H}_n=\bm{h}_n)>0$, $P_1( \bm{H}_n=\bm{h}_n) = P_2(\bm{H}_n=\bm{h}_n)>0$ and consequently
\begin{align}\label{eqn0}
P_1(\bm{V}_m=\bm{v}_m|\bm{H}_n=\bm{h}_n) =P_2(\bm{V}_m=\bm{v}_m |\bm{H}_n=\bm{h}_n).
\end{align}
Likewise, by the induction hypothesis applied to $(\bm{V}_{m-},\bm{H}_n)$ for $\bm{V}_{m-} = (V_2,\ldots,V_{m+1})$, we have
\begin{eqnarray}\label{eqn1}
 1- P_1(V_{m+1} = b | \bm{H}_n=\bm{h}_n) &=& P_1(V_{m+1} = a |\bm{H}_n=\bm{h}_n)\\
\nonumber  &=&  P_2(V_{m+1} = a | \bm{H}_n=\bm{h}_n)\\
\nonumber  & = &1-P_2(V_{m+1} = b | \bm{H}_n=\bm{h}_n)
\end{eqnarray}
for any $\bm{h}_n \in \{a,b\}^{n}$. Now fix  $\bm{v}_m \in \{a,b\}^{m}, \bm{h}_n \in \{a,b\}^{n}$ and let $\bm{v}_{m+1}= (v_{m+1},\bm{v}_m)$ for $v_{m+1}\in\{a,b\}$.  Then, by the conditional independence assumption,
$$
P_i( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n) = P_i( V_{m+1}=v_{m+1}| \bm{H}_n=\bm{h}_n) P_i( \bm{V}_{m}=v_{m}| \bm{H}_n=\bm{h}_n)
$$
for $i=1,2$,  so that we have $P_1( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n) = P_2( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n)$ by (\ref{eqn0})-(\ref{eqn1}). Consequently, it follows that
\begin{eqnarray*}
P_1( \bm{V}_{m+1}=\bm{v}_{m+1}, \bm{H}_n=\bm{h}_n) &=& P_1( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n) P_1(\bm{H}_n=\bm{h}_n)\\
& =& P_2( \bm{V}_{m+1}=\bm{v}_{m+1}| \bm{H}_n=\bm{h}_n)P_2(\bm{H}_n=\bm{h}_n) \\&=&P_2( \bm{V}_{m+1}=\bm{v}_{m+1}, \bm{H}_n=\bm{h}_n)
\end{eqnarray*}
by $P_1(\bm{H}_n=\bm{h}_n)=P_2(\bm{H}_n=\bm{h}_n)$. As $\bm{v}_{m+1}\in \{a,b\}^{m+1},\bm{h}_n \in \{a,b\}^{n}$ were arbitrary, this completes the proof of Lemma 1. $\Box$

\begin{lemma} Suppose the discrete random vector $(X,Y)$ has support $\{a,b\}\times \{a,b\}$, for some $a\neq b \in \mathbb{R}$,
under two probability distributions $P_1$ and $P_2$. Let $\E_i$ denote expectation under the distribution $P_i$, $i=1,2$. Then, the following are equivalent
\begin{enumerate}
\item $\E_1 X = \E_2 X$, $\E_1 Y = \E_2 Y$, $\E_1 X Y = \E_2 X Y$, where $\E_i$ denotes expectation under $P_i$, $i=1,2$.
\item $P_1(X=x,Y=y)=P_2(X=x,Y=y)$ for $x,y\in\{a,b\}$, i.e., $(X,Y)$ has the same distribution under $P_1$ and $P_2$.
\end{enumerate}
\end{lemma}

*Proof.* Denote the four probabilities by which the random vector $(X,Y)$ assume pairs $(b,b),(b,a),(a,b),(a,a)$, respectively, as $c_1,c_2,c_3,1-c_1-c_2-c_3$ under $P_1$ and $d_1,d_2,d_3,1-d_1-d_2-d_3$ under $P_2$. The conditions $\E_1 X = \E_2 X$, $\E_1 Y = \E_2 Y$, $\E_1 X Y = \E_2 X Y$ with $b\neq a$ imply that
$$
(c_1+c_2) = (d_1+d_2), \quad  (c_1+c_3) = (d_1+d_3) \quad (b-a) c_1 + a (2c_1+c_2+c_3) =  (b-a) d_1 + a (2d_1+d_2+d_3).
$$
As $(2c_1+c_2+c_3) = (2d_1+d_2+d_3)$ and $b\neq a$, we conclude that $c_1=d_1$, from which it follows that $c_2=d_2$ and $c_3=d_3$. Hence, claim 1 implies claim 2 in Lemma 2. Claim 2 also trivially implies claim 1. $\Box$

@@

Now suppose claim 3 holds. Pick $\ell \in \{1,\ldots,m\}$. Then, by the RBM$(m,n,\bm{\theta})$ probability structure and the iid Bernoulli assumption, we have the conditional probability
\begin{eqnarray*}
\frac{1}{2} &=& P\left(V_\ell = v_{\ell}\Big| \bigcap_{i=1 \atop i \neq \ell}^m \{V_i=v_i\} \cap \bigcap_{j=1}^n \{H_j=h_j\}\right) \\
&\equiv &\frac{\exp\left( v_\ell\theta_\ell +  v_\ell\sum_{j=1}^n  h_j \theta_{\ell j}\right)}{\exp\left(
\theta_\ell + \sum_{j=1}^n   h_j\theta_{\ell j} \right) +\exp\left(-\theta_\ell  -\sum_{j=1}^n  h_j\theta_{\ell j}\right) }
\end{eqnarray*}
for any subsets $\{v_i\}_{i=1}^m,\{h_j\}_{j=1}^n \subset\{\pm 1\}$. Equating conditional probabilities when $v_\ell=1$ or when $v_\ell=-1$, we have   $\theta_\ell = -  \sum_{j=1}^m  h_j \theta_{\ell j}$ for any given $\{h_j\}_{j=1}^n \subset \{\pm 1\}$. Taking all $h_j=1$ or all $h_j=-1$ for $1 \leq j \leq n$ yields $\theta_\ell=0$.  This result further implies $\theta_{\ell 1} = - \sum_{j=2}^n  h_j \theta_{\ell j}$ for any given $\{h_j\}_{j=2}^n \subset \{\pm 1\}$, so that $\theta_{\ell 1}=0$ must hold. Iterating this argument, sequentially for each $j=2,\ldots,m$, shows $\theta_{\ell j}=0$ for any $1 \leq j \leq n$ in addition to $\theta_{v_\ell}=0$. As $\ell\in\{1,\ldots,m\}$ was arbitrary, we have $\theta_{v_i}=0$ and $\theta_{i j}=0$ for any $1 \leq i \leq m$, $1 \leq j \leq n$.

### 2016-09-17 writing/draft.Rmd
formatting

```{r fitting}
source("helpers/functions.R")
load("data/sample_images.Rdata")

@@

#marginalized likelihood
load("data/fitted_models_trunc_marginal_thin.Rdata")
marginal_bad <- models_bad
marginal_good <- models_good

@@

#truncated normal
load("data/fitted_models_trunc_thin.Rdata")
trunc_bad <- models_bad
trunc_good <- models_good

@@

marginal_sample_good <- reshape_sample_distn(marginal_good)
trick_sample_good <- reshape_sample_distn(trick_good)
trunc_sample_good <- reshape_sample_distn(trunc_good)

marginal_sample_good %>% rename(marginal = prob) %>%
  left_join(trick_sample_good %>% ungroup() %>% select(-image_id) %>% rename(trick = prob)) %>%
  left_join(trunc_sample_good %>% ungroup() %>% select(-image_id) %>% rename(trunc = prob)) %>%
  ungroup() %>%
  mutate(image_id = paste0("image_", image_id)) %>% 
  gather(method, prob, trick, marginal, trunc) %>%
  spread(image_id, prob) -> all_statistics

>>>>

It should be noted that the truncated normal method (2) and the marginalized likelihood method (3) are drawing from the same stationary posterior distribution for images. The difference between these two methods is in how well the chains mix and how quickly they arrive at the target distribution. After a burn-in period of 50 iterations selected by inspecting the trace plots, we assess the issue of mixing in two ways. First, the autocorrelation functions (ACF) for each image are assessed at multiple lags for each method and plotted in figure \ref{fig:acf} with the truncated normal method in black and the marginalized likelihood method in red. As expected, ACF corresponding to the method that marginalizes out the hidden variables from the likelihood decreases to zero at a much faster rate, indicating better mixing for the chain. Secondly,

```{r models}
source("helpers/functions.R")
load("data/sample_images.Rdata")

@@

#marginalized likelihood
load("data/fitted_models_trunc_marginal_full.Rdata")
marginal_bad <- models_bad
marginal_good <- models_good

@@

#truncated normal
load("data/fitted_models_trunc_full.Rdata")
trunc_bad <- models_bad
trunc_good <- models_good

@@

marginal_sample_good <- reshape_sample_distn(marginal_good)
trick_sample_good <- reshape_sample_distn(trick_good)
trunc_sample_good <- reshape_sample_distn(trunc_good)
```

```{r acf, fig.cap="\\label{fig:acf}The autocorrelation functions (ACF) for each image assessed at multiple lags for each method with the truncated normal method in black and the marginalized likelihood method in red. As expected, ACF corresponding to the method that marginalizes out the hidden variables from the likelihood decreases to zero at a much faster rate, indicating better mixing for the chain."}
marginal_sample_good %>% 
  group_by(v1, v2, v3, v4, image_id) %>% 
  do(data.frame(acf = acf(.$prob, plot=FALSE)$acf,
     lag = acf(.$prob, plot=FALSE)$lag)) -> marginal_acfs

trunc_sample_good %>% 
  group_by(v1, v2, v3, v4, image_id) %>% 
  do(data.frame(acf = acf(.$prob, plot=FALSE)$acf,
                lag = acf(.$prob, plot=FALSE)$lag)) -> trunc_acfs

marginal_acfs %>% ungroup() %>% select(-image_id) %>% rename(marginal = acf) %>%
  left_join(trunc_acfs %>% rename(truncated = acf)) %>%
  ggplot() +
  geom_segment(aes(x=lag, xend=lag, y=0, yend=truncated), colour = "black") +
  geom_segment(aes(x=lag, xend=lag, y=0, yend=marginal), colour = "red") +
  facet_wrap(~image_id) +
  xlab("Lag") +
  ylab("ACF")

```{r posterior-dsn}
marginal_sample_good %>% rename(marginal = prob) %>%
  left_join(trick_sample_good %>% ungroup() %>% select(-image_id) %>% rename(trick = prob)) %>%
  left_join(trunc_sample_good %>% ungroup() %>% select(-image_id) %>% rename(trunc = prob)) %>%
  ungroup() %>%
  mutate(image_id = paste0("image_", image_id)) %>% 
  gather(method, prob, trick, marginal, trunc) %>%
  spread(image_id, prob) -> all_statistics
